Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-11-21 manuscript/main.tex

\author{%
    \parbox{\linewidth}{\centering
        Kevin S. Bonham,$^{1}$
        Guilherme Fahur Bottino,$^{1}$
        Shelley Hoeft McCann,$^{1}$
        Jennifer Beauchemin,$^{2}$
        Elizabeth Weisse,$^{3}$
        Fatoumata Barry,$^{2}$
        Rosa Cano Lorente,$^{2}$
        The RESONANCE Consortium$^{\ast}$,
        Curtis Huttenhower,$^{4-7}$
        Muriel Bruchhage,$^{3}$
        Viren D'Sa,$^{2}$
        Sean Deoni,$^{2}$
        Vanja Klepac-Ceraj$^{1\dagger}$
    }
\\
\normalsize{$^{1}$Department of Biological Sciences, Wellesley College, Wellesley, MA, USA}\\
\normalsize{$^{2}$Rhode Island Hospital, Providence, RI, USA}\\
\normalsize{$^{3}$Department of Psychology, University of Stavenger, Stavanger, Norway}\\
\normalsize{$^{4}$Department of Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{5}$Department of Immunology and Infectious Diseases, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{6}$Harvard Chan Microbiome in Public Health Center, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{7}$Associate Member, Broad Institute of MIT and Harvard}\\
\\
\normalsize{$^\ast$The RESONANCE consortium collaborators and affiliations are listed in the supplementary materials.}\\
\normalsize{$^\dagger$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}
}

@@

Shotgun metagenomic sequences were analyzed using the bioBakery suite of
computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
First, KneadData (v0.10.0) was used to perform quality
control of raw sequence reads, such as read trimming and removal of
reads matching a human genome reference. Next, MetaPhlAn (v3.1.0, using
database mpa\_v31\_CHOCOPhlAn\_201901) was used to generate taxonomic
profiles by aligning reads to a reference database of marker genes.
Finally, HUMAnN (v3.6.0) was used to functionally profile the
metagenomes.

@@

All data needed to evaluate the conclusions in the paper
are present in the paper and/or the Supplementary Materials.
Taxonomic and functional microbial profiles, as well as subject
demographics necessary for statistical analyses and machine learning are
available on the Data Dryad
\cite{bonhamDataDryad2023}.
Data processing, generation of summary statistics, and
generation of plots was performed using the julia programming language
and 3rd party packages
\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.
Information for replicating the package environment
as well as code for data analysis and figure generation, as well as scripts for automated
download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
and archived on Zenodo \cite{kevin_bonham_2023_7647510}.

>>>>

\author{%
    \parbox{\linewidth}{\centering
        Kevin S. Bonham,$^{1}$
        Guilherme Fahur Bottino,$^{1}$
        Shelley Hoeft McCann,$^{1}$
        Jennifer Beauchemin,$^{2}$
        Elizabeth Weisse,$^{3}$
        Fatoumata Barry,$^{2}$
        Rosa Cano Lorente,$^{2}$
        The RESONANCE Consortium$^{1\dagger}$,
        Curtis Huttenhower,$^{4-7}$
        Muriel Bruchhage,$^{3}$
        Viren D'Sa,$^{2}$
        Sean Deoni,$^{2}$
        Vanja Klepac-Ceraj$^{\ast}$
    }
\\
\normalsize{$^{1}$Department of Biological Sciences, Wellesley College, Wellesley, MA, USA}\\
\normalsize{$^{2}$Rhode Island Hospital, Providence, RI, USA}\\
\normalsize{$^{3}$Department of Psychology, University of Stavenger, Stavanger, Norway}\\
\normalsize{$^{4}$Department of Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{5}$Department of Immunology and Infectious Diseases, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{6}$Harvard Chan Microbiome in Public Health Center, Harvard T.H. Chan School of Public Health, Boston, MA, USA}\\
\normalsize{$^{7}$Associate Member, Broad Institute of MIT and Harvard}\\
\\
\normalsize{$^\dagger$The RESONANCE consortium authors and collaborators and affiliations are listed in the supplementary materials.}\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  vklepacc@wellesley.edu.}\\
}

@@

Shotgun metagenomic sequences were analyzed using the bioBakery suite of
computational tools \cite{beghiniIntegratingTaxonomicFunctional2021}.
First, KneadData (v0.10.0) was used to perform quality
control of raw sequence reads, such as read trimming and removal of
reads matching a human genome reference. Next, MetaPhlAn (v3.1.0, using
database mpa\_v31\_CHOCOPhlAn\_201901) was used to generate taxonomic
profiles by aligning reads to a reference database of marker genes.
Finally, HUMAnN (v3.6.0) was used to functionally profile the
metagenomes.
Statistical analyses, including linear regression
and feature set enrichment analysis were performed
using the julia programming language and third-party packages
\cite{bezansonJuliaFreshApproach2017,bonhamMicrobiomeJlBiobakeryUtils2021,danischMakieJlFlexible2021,blaomMLJJuliaPackage2020,ben_sadeghi_2022_7359268,dahua_lin_2023_7695673,douglas_bates_2023_7734970}.

@@

All data needed to evaluate the conclusions in the paper
are present in the paper and/or the Supplementary Materials.
Taxonomic and functional microbial profiles, as well as subject
demographics necessary for statistical analyses and machine learning are
available on the Data Dryad
\cite{bonhamDataDryad2023}.
Data processing, generation of summary statistics, and
generation of plots was performed using the julia programming language
and 3rd party packages (see Materials and Methods).
Information for replicating the package environment
as well as code for data analysis and figure generation, as well as scripts for automated
download of input files are available on GitHub at https://github.com/Klepac-Ceraj-Lab/Resonance,
and archived on Zenodo \cite{kevin_bonham_2023_7647510}.
